paper

### 2024-09-18 paper.qmd
.

# UPS1 Case Study 0

@@

## Run LimROTS

@@

## Run ROTS

@@

## Run limma

@@

## Run t-test

@@

## Comparison

@@

# UPS1 Case Study 1

## Run LimROTS

@@

## Run ANCOVA

@@

# UPS1 Case Study 2

@@

# BM21 Case study 3

## Run LimROTS

@@

# Load the dataset
load("../LimROTS paper/metabolomics/BM21.Case0.rda")

@@

@@

## Run ROTS

```{r}
# Load the dataset
load("../LimROTS paper/metabolomics/BM21.Case0.rda")
library(SummarizedExperiment)
library(doParallel)
library(LimROTS)
print(BM21.Case0)

@@

```{r}
length(which(rots.result$FDR < 0.05))
summary(rots.result$q_values)
plot(rots.result$q_values)
hist(rots.result$q_values)
hist(rots.result$p)
hist(rots.result$FDR)
```

## Run Limma

@@

load("../LimROTS paper/metabolomics/BM21.Case0.rda")

@@

@@

load("../LimROTS paper/metabolomics/BM21.Case1.rda")

print("BM21.Case1")
data <- assay(BM21.Case1)
formula.str <- "~ 0 + mixing.ratio"
meta.info <- data.frame(colData(BM21.Case1), check.names = FALSE, row.names = row.names(colData(BM21.Case1)))
meta.info$sample.id <- NULL
design.matrix <- model.matrix(formula(formula.str), data = meta.info)
colnames(design.matrix) <- make.names(colnames(design.matrix))
fit <- lmFit(data, design.matrix)
cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
fit2 <- contrasts.fit(fit, cont_matrix)
fit.ebayes <- eBayes(fit2, trend=T, robust=T)
limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")

>>>>

# UPS1-spiked DataSet

## UPS1 Case Study 0

@@

### Run LimROTS

@@

### Run ROTS

@@

### Run limma

@@

### Run t-test

@@

### Comparison

@@

## UPS1 Case Study 1

### Run LimROTS

@@

### Run ANCOVA

@@

## UPS1 Case Study 2

@@

# BM21 DataSet

To demonstrate the capability of LimROTS to also handle metabolomics data, the dataset was obtained from the Metabolomics Workbench (ST002454). Different matrix ratios of human plasma and vegetable juice were analyzed, and three different datasets were summarized into three *SummarizedExperiment* objects.

## BM21 Case study 3

This dataset contains six different mixing ratios: 1:1024, 1:256, 1:64, and their reverse, each with three technical replicates. The samples were grouped into high plasma ratio samples and low ratio samples. Peaks with more than 50% missing values were removed, and peaks with a variance less than 1 were excluded. Furthermore, only peaks quantified in samples with 100% plasma and 100% vegetable juice were selected. Therefore, we expect that all the peaks in this dataset represent true significant differences between the two mixing ratio groups.

### Run LimROTS

@@

# Load the dataset
data(BM21.Case0)

@@

# Set metadata and formula for LimROTS analysis
meta.info <- "mixing.ratio"
B <- 1000  # Number of bootstrap samples
K <- dim(BM21.Case0)[1]/4  # Set the value for K based on the data size
K <- floor(K) 
num_cores <- 30   # Number of cores for parallel processing
cluster <- makeCluster(num_cores)  # Create a parallel cluster
group.name <- "mixing.ratio"
formula.str <- "~ 0 + mixing.ratio"  # Formula for group comparison

@@

### Run ROTS

```{r}
# Load the dataset
library(SummarizedExperiment)
library(doParallel)
library(LimROTS)

data(BM21.Case0)
print(BM21.Case0)

@@

### Run Limma

@@

data(BM21.Case0)

@@

## BM21 Case study 4

@@

data(BM21.Case1)
print(BM21.Case1)
data <- assay(BM21.Case1)
formula.str <- "~ 0 + mixing.ratio"
meta.info <- data.frame(colData(BM21.Case1), check.names = FALSE, row.names = row.names(colData(BM21.Case1)))
meta.info$sample.id <- NULL
design.matrix <- model.matrix(formula(formula.str), data = meta.info)
colnames(design.matrix) <- make.names(colnames(design.matrix))
fit <- lmFit(data, design.matrix)
cont_matrix <- makeContrasts("mixing.ratio1-mixing.ratio2",  levels=design.matrix)
fit2 <- contrasts.fit(fit, cont_matrix)
fit.ebayes <- eBayes(fit2, trend=T, robust=T)
limma.result <- topTable(fit.ebayes, coef="mixing.ratio1-mixing.ratio2" , number = "Inf")
```
